---
title: "Text Mining on Yelp Reviews"
author: "Yiming Gao (yimingg2)"
date: "4/4/2018"
output: html_document
---

Some useful references:

- The Statistical Difference between 1-Star and 5-Star Reviews on Yelp: http://minimaxir.com/2014/09/one-star-five-stars/
- Analysis and Visualizations of Yelp Data using R and ggplot2: http://minimaxir.com/2015/12/lets-code-1/
- Sentiment Analysis of Restaurant Reviews: https://rpubs.com/yceron/155272 

The data set contains XX million reviews. It is formatted as by-line JSON: I wrote a pair of Python scripts to convert it to CSV for easy import into R.

The datasets are located here:

- Businesses Data: https://s3.us-east-2.amazonaws.com/578projectyelp/business.csv

Necessary packages:

```{r}
library(stringr)        #This package is used for string manipulation functions
library(dplyr)          #This package is used for data manipulation tasks
library(tidyr)          #This package is used for data manipulation tasks
library(data.table)     #This package is used to access the function fread which is a better/faster way to read large data
library(wordcloud)      #This package is used to generate word clouds
library(tm)             #This is a text mining package used in the process of generating word clouds
# library(RWeka)          # This package is used to generate Bigramws and Trigrams
library(ggplot2)        #This package is used for visualizations (chart/graph plotting functions)
library(ggmap)          #This package is used for map functions 
library(maps)           #This package is used for map functions
library(leaflet)        #This package is used for plotting maps
library(knitr)
```


### Read the data

```{r, message=FALSE, warning=FALSE}
data_business = fread("https://s3.us-east-2.amazonaws.com/578projectyelp/business.csv")

#Dimensions and attribute Names for the Business Data
dim(data_business)
```

