---
title: "CIFAR-10 Image Classification with CNN"
author: "Yiming Gao (yimingg2)"
date: "3/20/2018"
linestretch: 1.2
fontsize: 11pt
output:
  pdf_document: 
    toc: true
    number_sections: true
---
\newpage
# Data Set
## Introduction
The dataset I chose for HW2 is the CIFAR-10 dataset, which consists of 60000 32*32 color images in 10 classes, with 6000 images per class. There are 50000 training images and 10000 test images. (https://www.cs.toronto.edu/~kriz/cifar.html)

The dataset is divided into five training batches and one test batch, each with 10000 images. The test batch contains exactly 1000 randomly-selected images from each class. The training batches contain the remaining images in random order, but some training batches may contain more images from one class than another. Between them, the training batches contain exactly 5000 images from each class. 

Here are the classes in the dataset, as well as 10 random images from each:

\begin{center}
\includegraphics[width=12cm]{p1.PNG}
\end{center}

The classes are completely mutually exclusive. There is no overlap between automobiles and trucks. "Automobile" includes sedans, SUVs, things of that sort. "Truck" includes only big trucks. Neither includes pickup trucks.

## Model Architecture


## Goal& Approach
In this homework, I will use R to ingest data, train a model, and evaluate it on a test set. 

I build a dense CNN network for classifying the images into 10 classes, which are airplane, automobile, cat, deer, dog, frog, horse, ship, and truck. The usual method for training a network to perform N-way classification is multinomial logistic regression, known as softmax regression.

**Softmax** activation is used at the end which applies a softmax nonlinearity to the output of the network and calculates the **cross-entropy** loss between the probabilities calculated using the softmax activation function. For regularization, we also apply the usual weight decay losses to all learned variables.

### Data Processing

```{r, message=FALSE, warning=FALSE}
library(reticulate)
library(tensorflow)
library(keras)
```

Then we create training and test sets, and check the dimensions of them. From the output below, we know that we have 50000 training images and 10000 test images.

```{r}
cifar = dataset_cifar10()

# Training Data
train_x = cifar$train$x / 255
train_y = to_categorical(cifar$train$y, num_classes = 10)

# Test Data
test_x = cifar$test$x / 255
test_y = to_categorical(cifar$test$y, num_classes = 10)

# Check the dimentions
dim(train_x)

cat("No of training samples\t--", dim(train_x)[[1]], 
    "\tNo of test samples\t--", dim(test_x)[[1]])
```

### Model Definition

```{r}
model = keras_model_sequential()
```

